---
title: 'Project 5: AB Testing'
output:
  pdf_document: default
  html_notebook: default
---




# 0. Import Data

## Import Data
```{r}
# Load control, treatment, transactions df's
control <- read.csv('data/clean/control.csv')
treatment <- read.csv('data/clean/treatment.csv')
transactions <- read.csv('data/clean/transactions.csv')
```

## Check/Wrangle Data
```{r}
# Change transactions$invoice_date to datetime
# Load dependencies
library(lubridate)
# Change to datetime
transactions$invoice_date <- ymd(as.character(transactions$invoice_date))
# Check
str(transactions)

# Remove treatment units from control units
control <- subset(control, !(store_id %in% treatment$store_id))
```




# 1. Choose Target Variable
The goal is to increase gross margin per week, so the performance metric is 
weekly_gross_margin.




# 2. Choose Experimental Design
The data received includes descriptive data on both stores and transactions, 
so a matched pair design was used.




# 3. Choose Experimental and Control Variables

## Experimental Variable
The combination of updated restaurant menu and television advertising is the 
experimental variable because it was adjusted to determine its effect on 
the target variable, weekly gross margin.

## Control Variables

### List potential control variables we have data for
1. sq_ft
2. average monthly sales

### Wrangle data to obtain gross margins per store
```{r}
# Determine gross_margin

# Load dependencies
library(dplyr)

# gross margin per store
sum_transactions <- transactions %>% 
  group_by(store_id) %>%
  summarise(sum_gross_margin=sum(gross_margin))

# Join treatment/control df's to sum_gross_margin
treatment <- left_join(treatment, sum_transactions, by='store_id')
control <- left_join(control, sum_transactions, by='store_id')
```

### Check correlations between sum_gross_margin and potential control variables
```{r}
# Check correlations
cor(control[c('sum_gross_margin', 'sq_ft', 'avg_month_sales')])
```
Of the potential control variables of interest, sq_ft and avg_month_sales, only
avg_month_sales is correlated to the target variable, sum_gross_margin. As a 
result, avg_month_sales was selected as the only control variable.




# 4. Choose the Treatment and Control Units

## Treatment Units
The treatment units are the 10 stores in the treatment dataframe

## Control Units
Per project guidance, two control units from the control dataframe should be
matched to each treatment unit from the treatment dataframe.
```{r}
# Load dependencies
library(MatchIt)

# Include treatment column in treatment/control df's
treatment$treatment <- 1
control$treatment <- 0

# Divide Central and West Regions
central_units <- subset(rbind(control, treatment), region=='Central')
west_units <- subset(rbind(control, treatment), region=='West')

# Match the samples
# Central
matches_central <- matchit(treatment ~ avg_month_sales,
        data=central_units,
        method='nearest',
        ratio=2)
# West
matches_west <- matchit(treatment ~ avg_month_sales,
        data=west_units,
        method='nearest',
        ratio=2)

# Obtain matrix of matched units
matches_central <- matches_central$match.matrix
matches_west <- matches_west$match.matrix

# Create df of treatment and control store_id's
# Central
tu_central <- central_units[rownames(matches_central),]$store_id
cu1_central <- central_units[matches_central[,1],]$store_id
cu2_central <- central_units[matches_central[,2],]$store_id
match_matrix_central <- data.frame(tu_central, cu1_central, cu2_central)
names(match_matrix_central) <- c('Treatment Store', 'Control Store 1', 'Control Store 2')
# West
tu_west <- west_units[rownames(matches_west),]$store_id
cu1_west <- west_units[matches_west[,1],]$store_id
cu2_west <- west_units[matches_west[,2],]$store_id
match_matrix_west <- data.frame(tu_west, cu1_west, cu2_west)
names(match_matrix_west) <- c('Treatment Store', 'Control Store 1', 'Control Store 2')

# Combine match_matrix_central and match_matrix_west
match_matrix <- rbind(match_matrix_central, match_matrix_west)
# Display Treatment and Control Units
match_matrix
```




# 5. Clean and Prepare Data

## Helper Functions
```{r}
# Filters transaction data by date and calculates the weekly gross margin per store
# Inputs transaction data, start date, and end date
# Outputs a df with store_id, weekly_gross_margins, and id columns
wgm_by_dates <- function(df, date1, date2) {
  # Calculate weekly gross margins by date
  wgm_by_dates <- subset(df, invoice_date>=date1 & invoice_date<=date2) %>%
    group_by(invoice_week=cut(invoice_date, "week"), store_id) %>% 
    summarise(gross_margin=sum(gross_margin)) %>%
    group_by(store_id) %>%
    summarise(weekly_gross_margin=mean(gross_margin))
  # Add rownames as column to make joins easier
  wgm_by_dates$index <- as.numeric(row.names(wgm_by_dates))
  return(wgm_by_dates)
}
```

## Comparative Dates
```{r}
# Filter data by test dates (2015-04-29 to 2015-07-21)
# and calculate weekly_gross_margin
wgm_comparative_dates <- wgm_by_dates(transactions, '2015-04-29', '2015-07-21')

# Create dataframe of weekly_gross_margins for treatment/control units

# TREATMENT
wgm_comparative_treatment <- left_join(match_matrix, wgm_comparative_dates, by=c('Treatment Store'='store_id'))$weekly_gross_margin

# CONTROL
wgm_comparative_control1 <- left_join(match_matrix, wgm_comparative_dates, by=c('Control Store 1'='store_id'))$weekly_gross_margin
wgm_comparative_control2 <- left_join(match_matrix, wgm_comparative_dates, by=c('Control Store 2'='store_id'))$weekly_gross_margin
wgm_comparative_control <- rowMeans(data.frame(wgm_comparative_control1, wgm_comparative_control2))
# Create dataframe of treatment and control weekly_gross_margin
wgm_comparative <- data.frame(wgm_comparative_treatment, wgm_comparative_control)
names(wgm_comparative) <- c('Treatment', 'Control')
```

## Test Dates
```{r}
# Filter data by test dates (2016-04-29 to 2016-07-21)
# and calculate weekly_gross_margin
wgm_test_dates <- wgm_by_dates(transactions, '2016-04-29', '2016-07-21')

# Create dataframe of weekly_gross_margins for treatment/control units

# TREATMENT
wgm_test_treatment <- left_join(match_matrix, wgm_test_dates, by=c('Treatment Store'='store_id'))$weekly_gross_margin

# CONTROL
wgm_test_control1 <- left_join(match_matrix, wgm_test_dates, by=c('Control Store 1'='store_id'))$weekly_gross_margin
wgm_test_control2 <- left_join(match_matrix, wgm_test_dates, by=c('Control Store 2'='store_id'))$weekly_gross_margin
wgm_test_control <- rowMeans(data.frame(wgm_test_control1, wgm_test_control2))
# Create dataframe of treatment and control weekly_gross_margin
wgm_test <- data.frame(wgm_test_treatment, wgm_test_control)
names(wgm_test) <- c('Treatment', 'Control')
```




# 6. Calculate Lift

```{r}
# Create df of increase in wgm by control, increase in wgm by treatment
lift_control <- wgm_test$Control/wgm_comparative$Control
lift_treatment <- wgm_test$Treatment/wgm_comparative$Treatment
lift <- data.frame(lift_control, lift_treatment)

# Calculate increase in weekly_gross_margin of treatment relative to control
lift_overall <- mean(lift$lift_treatment)/mean(lift$lift_control)
lift_west <- mean(lift$lift_treatment[1:5])/mean(lift$lift_control[1:5])
lift_central <- mean(lift$lift_treatment[6:10])/mean(lift$lift_control[6:10])
lift_overall
lift_west
lift_central
```
The calculated overall lift is approximately 46%, while the calculated lift for
west region stores is approximately 43% and the calculated lift for central
region stores is approximately 49%.




# 7. Calculate Statistical Significance

```{r}
# Paired T-Test because the test conducted was matched pair
# Overall lift
t.test(lift$lift_treatment, lift$lift_control, paired=TRUE)
# Central lift
t.test(lift$lift_treatment[1:5], lift$lift_control[1:5], paired=TRUE)
# West lift
t.test(lift$lift_treatment[6:10], lift$lift_control[6:10], paired=TRUE)
```
With a p-value of 0.0001156, there is sufficient evidence to reject the null 
hypothesis, that there is no difference in weekly gross margins between the 
control and treatment groups, in favor of the alternative hypothesis, that there 
is a difference in weekly gross margins between the control and treatment 
groups. The t-test estimated lift is approximately 44% for all stores, 42% for
west region stores and 46% for central region stores.



# 8. Summary

## Planning the Analysis
Round Roasters, a coffee restaurant in the US, conducted a market test to 
determine the impacts of introducting a new menu and television marketing 
campaign. Because Round Roasters was interested in profitability and buying
coffee follows a weekly cycle for customers, data was aggregated weekly and
weekly gross margin was used as the target metric. Because Round Roasters 
collected data on their stores and transactions, a matched pair analysis was 
used. This matched pair analysis was conducted using the test period (April 29, 
2016 to July 21, 2016) and the comparative period (April 29, 2015 to July 21, 
2015).

## Matching Treatent and Control Units

### Determining Control Units
Two potential control variables were chosen for their logical validity and
data availability: 
1. square footage 
2. average monthly sales

Per the correlation matrix below, average monthly sales correlates with gross
margin, with a correlation coefficient of 0.995, while square footage does not
correlate with gross margin, with a correlation coefficient of -0.029. Thus,
average monthly sales was the only control variable used in the analysis.
```{r echo=FALSE}
cor(control[c('sum_gross_margin', 'sq_ft', 'avg_month_sales')])
```


### Matching Treatment and Control Units

The following control units were matched to their respective treatment units, 
controlling for weekly sales and region of each store.
```{r echo=FALSE}
match_matrix
```

## Analysis

A paired t-test was conducted between the increase in weekly gross sales between
the test period (April 29, 2016 to July 21, 2016) and the comparative period 
(April 29, 2015 to July 21, 2015) for the treatment units and average of the 
control units. With a p-value of 0.0001156, there is sufficient evidence to 
reject the null hypothesis, that there is no difference in weekly gross margin 
between the control and treatment groups, in favor of the alternative 
hypothesis, that there is a difference in weekly gross margin between the 
control and treatment groups. Per the bar graph of lift for each 
treatment-control unit pair below, the increase in incremental lift between 
treatment and control units is 44% for all stores. Because the increase in 
incremental lift for the new menu is statistically significant in both central 
region stores, at 46%, and west region stores, at 42%, it should be implemented 
in both central and west region stores.

```{r echo=FALSE}
# Load dependencies
library(ggplot2)

# Visualize lift differences between treatment and control units
# Add column for unit
lift$control_number <- rownames(lift)
ggplot(aes(x=control_number, y=lift_treatment, color='Treatment'), data=lift) +
  geom_col() +
  geom_col(aes(y=lift_control, color='Control')) +
  xlab('Treatment-Control Pair') +
  ylab('Lift') +
  ggtitle('Lift for Treatment and Control Units')
```

