---
title: "AB Test - Summary"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Summary of AB Analysis

Round Roasters, a coffee restaurant in the US, conducted a market test to 
determine the impacts of introducting a new menu and television marketing 
campaign. Because Round Roasters was interested in profitability, weekly gross
margin was used as the target metric. Because Round Roasters collected data on
their stores and transactions, a matched pair analysis was used. After testing
potential control variables for their logical validity, distribution across 
potential control units, correlation to the target metric, and covariance, 
weekly_sales and region were chosen. 2 control units were then matched to each
treatment unit, with the following results:
```{r echo=FALSE}
matched_units
```
Next, the Treatment and Control Units were compared to determine the new menu
and marketing campaign's impact on profitability.

# Conclusions

A paired t-test was conducted between the increase in weekly gross sales between
the test period (April 29, 2016 to July 21, 2016) and the comparative period 
(April 29, 2015 to July 21, 2015) for the treatment and average of the control
units. With a p-value of 5.234e-05, there is sufficient evidence to reject the 
null hypothesis, that there is no difference in weekly gross margin between the 
control and treatment groups, in favor of the alternative hypothesis, that there 
is a difference in weekly gross margins between the control and treatment 
groups. The estimated incremental lift is approximately 39%.
```{r echo=FALSE}
# Paired T-Test because the test conducted was matched pair
t.test(lift$lift_treatment, lift$lift_control, paired=TRUE)
```